Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-02-04 ms.tex
More changes to the detection of black holes appendix...

>>>>

\subsection{X-ray Diagnostics of SMBH Occupation}
\label{sec:occ}
The free-free emission from steady state gas flows in low mass
galactic nuclei may serve as useful diagnostics of the presence or
absence of a SMBH.  We focus primarily on low-mass galaxies in this
subsection, both because this is the portion of parameter space in
which gas inflow will generally proceed in a steady-state fashion, and
because of recent observations suggesting that a nontrivial fraction
of these galaxies may lack nuclear SMBHs.  If this absence is
confirmed, it could have implications for galaxy evolution, SMBH
recoil distributions, and SMBH seed formation models.

We can approximate the gas temperature as

\begin{equation}
% T \approx \frac{\mu m_{\rm p} (\gamma-1)}{2\gamma k_{\rm B}}
% \left(\tilde{v}_{\rm w}^2 + GM_\bullet/r \right).
 \kb T \approx 2.1\times 10^{-25} \left(\tilde{v}_{\rm w}^2 + GM_\bullet/r
 \right) {\rm [cgs]}.
\end{equation}

In the absence of a central SMBH, we can expect to see an exponential
cutoff in the free-free X-ray spectrum at energies above

\begin{equation}
E_{\rm cut} = 0.32~{\rm keV} v_{500}^2,
\end{equation}

assuming $\gamma = 5/3$.  This cutoff will therefore be located in the
portion of the soft X-ray covered by {\it Chandra}'s ACIS
spectrometer.  By writing the free-free emissivity as

\begin{equation} 
\varepsilon_{\nu}^{\rm ff} = 6.8\times10^{-38}
\frac{\rm erg}{\rm s~cm^3~Hz} Z^2n_{\rm e}n_{\rm
i}T^{-1/2}\bar{g}_{\rm ff}\exp\left(\frac{-h\nu}{k_{\rm B}T}\right),
\end{equation}

where $Z$ is the atomic number of the ions, $n_{\rm e}$ and $n_{\rm
  i}$ are respectively the electron and ion number density, and
$\bar{g}_{\rm ff}$ is the Gaunt factor.  In the following we set
$n_{\rm e}=Y_{\rm e}n_{\rm i}$, and approximate
\begin{equation}
n_{\rm i} \approx \frac{\dot{M}}{(4\pi/3) \rs^2 \mu \mp}\sqrt{ \frac{r_{\rm s}}{ 2 G \Mbh} } \left( \frac{r}{\rs} \right)^{-n}.
\end{equation}

We now write the total spectral luminosity density seen in unresolved
X-ray observations above the cutoff energy as

\begin{align}
 L_{\nu}^{\rm ff} =& \int_0^{r_{\rm s}} 4\pi r^2 \varepsilon_{\nu}^{\rm ff}{\rm d}r \\
\approx & 2.8\times 10^{16} \, {\rm ergs \, s^{-1}
  \, Hz^{-1}} \left(\frac{\dot{M}}{10^{-5}M_\odot/{\rm
      year}} \right)^2 \left(\frac{\rs}{10^{-2}~{\rm pc}} \right)^{-1}
   \notag \\
& \times \left(\frac{h \nu}{0.5~{\rm keV}}\right)^{-3/2} \bar{Z}^2 \bar{g}_{\rm ff} Y_{\rm e} \notag \\
\approx & 4.6\times 10^{15} {\rm ergs~s^{-1}~Hz^{-1}}
\left(\frac{\dot{M}}{10^{-5}M_\odot/{\rm year}} \right)^2
\left(\frac{\Mbh}{10^{6}~{\Msun}} \right)^{-1} \notag\\&\times
\left(\frac{h \nu}{0.5~{\rm keV}}\right)^{-3/2} v_{500}^2 \bar{Z}^2 \bar{g}_{\rm ff} Y_{\rm e} 
\end{align}

The second line of this equation makes the approximation that
$GM_\bullet/r$ dominates $\tilde{v}_{\rm w}^2$ throughout the
integrand, which will be true for energies modestly higher than
$E_{\rm cut}$.  The lower (upper) endpoint for this integral are
arbitrary so long as it is somewhat less (greater) than the radius at
which the observed frequency $\nu$ sees its exponential free-free
cutoff. We have also set $n=1$.

In general, these ``cut-off'' X-ray luminosities are quite small, and
even {\it Chandra} would have little hope of detecting them in any but
the nearest dwarf galaxies (e.g. the Magellanic clouds).  However, if
source confusion is not present, this technique could test for the
presence of IMBHs in very nearby dwarf galaxies, or more speculatively
Galactic star clusters.
